\section{Related Work}

\paragraph{Jacobian representations for robot action.}
Neural Jacobian Fields learn visual Jacobian fields that map desired visual changes to robot actions across diverse robots \citep{li2024jacobian}. VERA uses a video planner with a Jacobian-based inverse dynamics module to produce generalist robot policies \citep{li2026vera}. Our work does not improve NJF or reproduce VERA-scale robot experiments. It uses the same kind of local action-sensitivity object for a different purpose: diagnosing whether a generated transition is locally executable and repairing trajectories at inference time.

\paragraph{Executable video and world-action verification.}
EVA aligns video world models with executable robot actions through inverse-dynamics rewards \citep{wang2026eva}. World Action Verifier studies self-improving world models through forward-inverse asymmetry \citep{liu2026wav}. These works are the closest novelty boundary. Our residual is not a learned verifier loop or a large-video-model finetuning objective. It is a low-compute tangent-space certificate that can be computed from analytic, finite-difference, or small learned action maps.

\paragraph{Energy-based planning and compositional inference.}
Energy-based planning frames control as inference over composed objectives \citep{du2019planning}. LEAP performs iterative energy minimization with sequence models \citep{chen2023leap}, and compositional generative modeling studies how separate energy factors can be combined at inference time \citep{du2024compositional}. We use actionability as one factor in such a composed trajectory energy, rather than as a decorative name.

\paragraph{Queryable scene and world representations.}
Neural scene representations for visuomotor control demonstrate that learned representations can support robot control queries rather than only rendering \citep{li20213d}. RoboDreamer and Large Video Planner push world-model imagination toward robot planning \citep{zhou2024robodreamer,chen2025lvp}. Our claim is narrower: a representation for embodied planning should expose local actionability, not only visual prediction or geometric state.

\paragraph{Failure modes in passive learning.}
Visual foresight and model-predictive control showed that learned video prediction can be useful for manipulation \citep{finn2017foresight}. Causal confusion and observational overfitting show that passive correlations can mislead policies or value estimates \citep{dehaan2019causal,song2019observational}. Actionability is a related embodied failure mode: a passive future can be plausible but physically unreachable for the current body.
